\bfseries Convolution-based functions on tensor fields and tensor images

### 2022-08-23 paper/main.tex
furniture re-arranged

We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total). Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel). Sometimes we will consider the image to be in a $d$-torus.

\section{Linear functions of geometric images}

%Similarly, each linear geometric convolution operator $C$ will be a pixelized cubical image patch with $M$ pixels (with $M$ odd) in each dimension ($M^d$ pixels total) with a tensor or pseudotensor in each pixel, which acts as a linear pixel weight.

For some of the expressions below we will use Ricci and Levi-Civita summation notation (or Einstein summation notation).
In this notation, tensor products are written in component form, and repeated indices are summed over.
In this notation, the product of two 2-tensors (represented as two $d\times d$ matrices $A$ and $B$) is written as
\begin{equation}
    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} ~,
\end{equation}
where $[A]_{i,k}$ (for example) is the $i,k$ element of matrix $A$, the repeated index $k$ is implicitly summed over the integers from 1 to $d$.
This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice).

One consequence of the above definitions is that 
$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor iff (in summation notation)
$[g\cdot T]_{i_1,\ldots, i_k} = [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}$ for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
For example, a 2-tensor is defined by its transformation property
$[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
which, in normal matrix notation, is written as
$g\cdot T = M(g)\,T\,M(g)^\top$.

@@

\begin{definition}[Kronecker contractions of tensors]
Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$. We define the $(\alpha,\beta)$ Kronecker contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$. In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$s are placed in the positions $\alpha$ and $\beta$:
\begin{equation}
[a^{(\alpha,\beta)}]_{i_1,\ldots, i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, i_{\alpha-1}, x, i_{\alpha+1}, \ldots, i_{\beta-1}, x, i_{\beta+1}. \ldots, i_k}
\end{equation}
There are $\choose{k}{2}$ different index choices for this contraction, corresponding to all the possible index pairs among $k$ original indices.
\end{definition}

\begin{definition}[Levi-Civita contractions of tensors]
Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq d-1$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$. We define the Levi-Civita contraction of $a$ to be the $(k-d+2)$-$(p+1)$-tensor resulting from taking the outer product of 
$a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol. There are $(d-1)!\,\choose{k}{d-1}$ different index choices for this contraction corresponding to an ordered list of $d-1$ indices chosen from among $k$ choices.
\end{definition}

\begin{definition}[permutations of tensor indices]
Given a $k$-tensor or $k$-pseudotensor $T\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $T^\sigma$ the result of permuting the indices of $T$ according to $\sigma$:
\begin{equation}
    [T^\sigma]_{i_1, \ldots, i_k} := [T]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
\end{equation}
\end{definition}

\begin{definition}[$k$-tensor image]
We say that $A$ is a $k$-tensor image if $A$ is a function $A: [N]^d \to (\mathbb R^d)^{\otimes k}$ such that $A(\bar \i)$ is a $k$-tensor for all $\bar \i \in [N]^d$, and $[N]=\{0,\ldots, N-1\}$.
By a similar definition there are also $k$-pseudotensor images.
We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
\end{definition}

\section{Linear functions of geometric images}\label{sec:convolution}

In this \sectionname{} we define the linear operations that can be performed on a geometric image such that the outputs of these operations are also geometric images.

\begin{definition}[geometric convolution]
Let $A$ be a $k$-tensor image on the $d$-torus with side length $N$.
Let $C$ be a $k'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
The geometric convolution $A\ast C$ is a $k+k'$-tensor image such that
\begin{equation}
    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
\end{equation}
where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
If $A$ is a $k$-pseudotensor image and $C$ is a $k'$-pseudotensor filter, then $A\ast C$ is also a $k+k'$-tensor.
If $A$ is $k$-tensor and $C$ is $k'$-pseudotensor, or if $A$ is $k$-pseudotensor and $C$ is $k'$-tensor, then $A\ast C$ is a $k+k'$-pseudotensor.
\end{definition}

Geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor inputs and outputs.

The linear geometric convolution operation can be followed by contractions with Kronecker deltas and Levi-Civita symbols to make lower-$k$, but still linear, objects.
The Kronecker delta is the object with 2 indices such that if the two indices are identical the value is $1$ and the value is $0$ otherwise.
The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-tensor.
Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired. SOLE DO WE NEED TO SAY MORE HERE?
Contraction of a $k$-pseudotensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-tensor.

For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.

@@

@@

@@

\appendix

\section{All that follows needs to move up above}

@@

SOLE: WHERE SHOULD THIS GO?
Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY? IF NOT, IT SHOULD BE DEFINED. IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
Here are the linear operations (linear in the input image) that are permitted):
\begin{itemize}
\item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
  Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
\item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
  Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
  Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions. (SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
\end{itemize}

\section{Nonlinear equivariant local geometric operators}\label{sec:nonlinear}

\section{Expressive power of this architecture}
We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).

\section{Construction of equivariant geometric convolution filters}
In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
Here we show how to find all such filters by a group-averaging technique.

@@

@@

\section{Implementation notes}
\begin{itemize}
    \item find the invariant geometric filters
    \item geometric convolution operator
    \item pixel-wise outer product
    \item contraction operator, enumeration of all contractions
    \item output index permutation, enumeration of all permutations
    \item enumeration of all monomials up to a fix order and degree 
    \item remove repeated objects (SVD or just lstsq)?
    \item linear combinations 
    \item loss function
    \item optimization
\end{itemize}

\section{Discussion and comments}

>>>>

\begin{definition}[summation notation]
In this notation, tensor products are written in component form, and repeated indices are summed over.
In this notation, the product of two 2-tensors (represented as two $d\times d$ matrices $A$ and $B$) is written as
\begin{equation}
    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} ~,
\end{equation}
where $[A]_{i,k}$ (for example) is the $i,k$ element of matrix $A$, the repeated index $k$ is implicitly summed over the integers from 1 to $d$.
This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice). 
\end{definition}

One consequence of the above definitions is that 
$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-$p$-tensor if and only if (in summation notation)
\begin{align}
    [g\cdot T]_{i_1,\ldots, i_k} = \det(M(g))^p\, [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}
\end{align} for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
For example, a 2-0-tensor is defined by its transformation property
$[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
which, in normal matrix notation, is written as
$g\cdot T = M(g)\,T\,M(g)^\top$.

We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total). Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel). Sometimes we will consider the image to be in a $d$-torus. 
Let $\mathcal T_{d,k,p}$ the set of $k$-$p$-tensors in $\mathbb R^d$. We define the geometric images as follows.
\begin{definition}[geometric image]
A geometric image is a function $A:[N]^d \to \mathcal T_{d,k,p}$, where $[N]=\{0,1,\ldots, N-1\}$. We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
\end{definition}

\section{Generalized convolutions on geometric images}\label{sec:convolution}

In this \sectionname{} we generalize the notion of convolution to take geometric images as inputs and return geometric images as outputs.
The idea is that the (presumably large) geometric image (containing $k$-$p$-tensors) is convolved with a (presumably small) geometric filter (also a geometric image, but containing $k'$-$p'$-tensors) to produce an output image that contains $(k+k')$-$(p+p')$-tensors, where each pixel is a sum of outer products.
These outer products can be contracted down to lower-$k$ tensors using Kronecker or Levi-Civita contractions.

@@

If the next definition is going to use ``the torus'' then we should say a few words about that up-front, or above in the image definition.

\begin{definition}[geometric convolution]
Let $A$ be a $k$-$p$-tensor image on the $d$-torus with side length $N$.
Let $C$ be a $k'$-$p'$-tensor image (filter) on $[-m, m]^d$ where $2m+1<N$.
The geometric convolution $A\ast C$ is a $(k+k')$-$(p+p')$-tensor image such that
\begin{equation}
    (A\ast C)(\bar\i) = \sum_{\bar a\in[-m, m]^d} A(\bar\i - \bar a)\otimes C(\bar a) ~,
\end{equation}
where $\bar\i - \bar a$ is the translation of $\bar\i$ by $\bar a$ on the $d$-torus pixel grid $(\mathbb Z / N\mathbb Z)^d$.
\end{definition}
This definition is on the torus. If you are not in the torus just pad with zero tensors of the corresponding order and parity.

Convolutions can be followed by Kronecker contractions, Levi-Civita contractions and index permutations.
The Kronecker delta is the object with two indices such that it has the value $+1$ when the two indices have the same value, and $0$ otherwise.
Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-tensor.
Contraction of a $k$-$p$-tensor (with $k\geq 2$) with one application of the Kronecker delta returns a $(k-2)$-$p$-pseudotensor, because the contraction sums out one pair of indices.
The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices (where $d$ is the spatial dimension) such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
Contraction of a $k$-$p$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-$(p+1)$-pseudotensor, because the contraction sums over $d$ pairs of indices, and it changes the parity from even to odd or odd to even.
For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a pseudovector.

\begin{definition}[Kronecker contractions of tensors]
Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq 2$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$. We define the $(\alpha,\beta)$ Kronecker contraction of $a$ to be the $k-2$ tensor resulting from summing jointly over the indices $i_\alpha$ and $i_\beta$ where $\alpha < \beta \in\{1,\ldots, k\}$. In Einstein summation notation it would be $a_{i_1, \ldots, x, \ldots , x, \ldots, i_k}$ where the $x$s are placed in the positions $\alpha$ and $\beta$:
\begin{equation}
[a^{(\alpha,\beta)}]_{i_1,\ldots, i_{\alpha-1}, i_{\alpha+1}, \ldots, i_{\beta-1}, i_{\beta+1} , \ldots, i_k} := \sum_{x=1}^d [a]_{i_1, \ldots, i_{\alpha-1}, x, i_{\alpha+1}, \ldots, i_{\beta-1}, x, i_{\beta+1}. \ldots, i_k}
\end{equation}
\end{definition}
There are $\choose{k}{2}$ different index choices for this contraction, corresponding to all the possible index pairs among $k$ original indices.

\begin{definition}[Levi-Civita contractions of tensors]
Let $a$ be a $k$-$p$-tensor in $\mathbb R^d$ with $k\geq d-1$ such that $[a]_{i_1,\ldots i_k}\in \mathbb R$ for all $i_s = 1,\ldots, d$, and $s=1,\ldots, k$. We define the Levi-Civita contraction of $a$ to be the $(k-d+2)$-$(p+1)$-tensor resulting from taking the outer product of 
$a$ with the $d$-index Levi-Civita symbol $\epsilon_{i_1,i_2,\ldots,i_d}$ and then performing $d-1$ Kronecker contractions in which the first index is chosen from the (remaining) indices of $a$ and the second index is chosen from the (remaining) indices of the Levi-Civita symbol.
\end{definition}
There are $(d-1)!\,\choose{k}{d-1}$ different index choices for this contraction corresponding to an ordered list of $d-1$ indices chosen from among $k$ choices.

\begin{definition}[permutations of tensor indices]
Given a $k$-$p$-tensor $a\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $a^\sigma$ the result of permuting the indices of $a$ according to $\sigma$:
\begin{equation}
    [a^\sigma]_{i_1, \ldots, i_k} := [a]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
\end{equation}
\end{definition}

Define POOLING operation?

The following is not correct -- doesn't include permutations or pooling, and doesn't get the combinatorics right.

Now imagine that we want to find \emph{all} linear equivariant functions that go from an input $k$-tensor image of parity $p$ (SOLE: DO WE USE THIS CONCEPT OF PARITY PREVIOUSLY? IF NOT, IT SHOULD BE DEFINED. IT MIGHT SIMPLIFY OUR SHIT.) to an output $k'$-tensor image of parity $p'$, such that the input and output images are of the same size (both $N^d$).
Here are the linear operations (linear in the input image) that are permitted):
\begin{itemize}
\item Choose an integer $\ell$ in the range $0\leq\ell\leq k$ such that $(k'-k+2\,\ell)>0$.
  Convolve the input $k$-tensor image with a $(k'-k+2\,\ell)$-tensor invariant filter of parity $p\,p'$.
  Contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing at least one index from the indices of the input image.
  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions.
\item Choose an integer $\ell$ in the range $0\leq\ell\leq k-d+1$ such that $(k'-k+d-2+2\,\ell)>0$.
  Convolve the input $k$-tensor image with a $(k'-k+d-2+2\,\ell)$-tensor invariant filter of parity $-p\,p'$.
  Contract the resulting $(k'+d-2+2\,\ell)$-tensor image (of parity $-p'$) with the Levi--Civita symbol.
  There are $$\left[\choose{k'+d-2+2\,\ell}{d-1}-\choose{k'-k+d-2+2\,\ell}{d-1}\right]$$ choices for this Levi--Civita contraction.
  Then contract the resulting $(k'+2\,\ell)$-tensor image (of parity $p'$) with the Kronecker delta $\ell$ times, always choosing one index from the indices of either the input image or of the Levi--Civita symbol.
  There are $$\prod_{i=1}^{\ell}\frac{1}{\ell!}\,\left[\choose{k'+2\,i}{2}-\choose{k'-k+2\,i}{2}\right]$$ possible choices for these $\ell$ Kronecker contractions. (SOLE: IS THIS CONSISTENT WITH LANGUAGE ABOVE?)
\end{itemize}

other things that may be useful: eigenvalues

is it obvious how this contractions and permutations apply to the entire image? the same permutation, contraction should be applied to all pixels of the image.

@@

The convolution and contraction operations above defines a large class of linear functions from geometric images to geometric images. Non-linear functions can be constructed as products.

Algorithm: \newline
Construct all linear functions using the constructions from previous section. \newline
Construct all pixel-wise products across them. \newline
All linear functions of all pixel-wise products. \newline
That gives everything to second order. \newline

Assembling the parts into a general, linear function.

@@

Assemble all the parts into an algorithm for making all possible monomials.

@@

Define equivariance.

Prove translation invariance.

Prove thing about invariant filters.

@@

\subsection{Construction of equivariant geometric convolution filters}
In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
Here we show how to find all such filters by a group-averaging technique.

@@

Counting arguments.

Comments on locality.

We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).

How do we make the invariant filters?

How do we make the linear functions?

What do they look like?

Etc.

An example?

@@

\begin{itemize}
    \item find the invariant geometric filters
    \item geometric convolution operator
    \item pixel-wise outer product
    \item contraction operator, enumeration of all contractions
    \item output index permutation, enumeration of all permutations
    \item enumeration of all monomials up to a fix order and degree 
    \item remove repeated objects (SVD or just lstsq)?
    \item linear combinations 
    \item loss function
    \item optimization
\end{itemize}

One discussion point: you might be able to get back to a continuous symmetry group if you consider the image to be a VIEW of a large, continuous tensor field...?
